\chapter{Related Work}
\label{chap:related}

There are a number of related efforts that also seek to fulfill one or
more of the goals of this project as outlined in
Section~\ref{chap:intro:goals}. This chapter describes some of the
more pertinent of these efforts.

\section{Trust, Threat, and Security Models}
\label{chap:related:models}

Security researchers have developed a number of trust, threat, and
security models for a range of computing systems. Some of these models
are concerned merely with the technical security of the system. Others
expand to account for the many proclivities of human behavior that
have bearings on the security and privacy of computing systems. In all
cases, such models seek to answer the questions:

\begin{packed_item}
\item ``On what assumptions does the security of a given system rest?''
\item ``In what manner can those assumptions be violated?''
\item ``What is the effect of violating such assumptions?''
\end{packed_item}

Security models are a critical part of designing any security
system. The security of a given system is only as good as its weakest
link. Security models provide an analyses of the various links within
and surrounding any given security or privacy enhancing system.  Such
links that must be considered when determining the overall guarantees
provided by a given system as well as the exposure the system to
various threats.

\cite{flowerday2006}~provides an analyses of how trust and related
human controls must be considered in the design of any information
security system. It builds on work outside the traditional computer
science space related to theories of trust and privacy in government,
law, and society~\cite{camp2003}. Such works lay the foundation for
trust analysis in computing systems and tie the concept of trust in
computing systems to wider theories of trust and privacy as
fundamental concepts.

Beyond trust analyses, there are a number of more general computer
security models and taxonomies that focus on traditional technical and
operational risks and mitigations. \cite{abbas2005}~discuses security
and threat models for Internet-connected services.
\cite{tsipenyuk2005}~focuses on common failures in application-level
security. \cite{firesmith2005}~dives into a more abstract analysis of
safety and security in computing systems and provides a framework for
analyzing software security risks, potential harms, and security
requirements. Finally, \cite{cebula2010}~takes a comprehensive look at
operational cyber-security risks across technical, human, process, and
external factors.

While all of these efforts provided a basis for trust, threat, and
security analysis of computing systems, they do not dive into the
specific intricacies of the third party trust requirements inherent in
cloud computing systems. This is a deficiency Chapter~\ref{chap:trust}
seeks to correct.

\section{Minimizing Third Party Trust}
\label{chap:related:minimize}

Uneasiness with having to trust third parties has led to a number of
efforts to reduce, limit, and control such trust. Many of these
efforts aim to leverage cryptographic primitives as an alternative to
a trusted third party (TPP). As mentioned in
Section~\ref{chap:background:crypto}, cryptographically-based systems
generally require little to no trust in external systems or parties,
only in the underlying math (and associated implementations). Other
efforts use information theory primitives such as secret sharing
schemes (also discussed in Section~\ref{chap:background:crypto}) to
try to limit how much damage any single TPP can do. Such efforts
generally require multiple third parties to collude to accomplish most
attacks. In many cases, these efforts also explore the trade-offs
between the convenience and capabilities that trusted third party
systems can provide and the security risks of requiring one or more
trusted third parties.

\subsection{Cryptographic Access Control}

The primary limitation of all of the access control models mentioned
in Section~\ref{chap:background:ac} is their reliance on a trusted
arbiter for enforcement: generally this trusted arbiter is the
operating system or third party underlying a given access control
scheme. This means that the security of these access control systems
is only as good as the security of the system or third party enforcing
them. Thus, if the underlying OS or third party is compromised, the
access control system fails. Likewise, anyone in control of the
underlying OS or third party (e.g. an administrator) automatically
gains full control over the access control system and the ability to
bypass it\footnote{E.g. as in the case of Edward Snowden's collection
  of large troves of secret NSA data from a facility where he acted as
  a systems administrator.}. This is an acceptable limitations in many
situations, especially those based on centrally managed systems with
existing physical and administrative safeguards in place. But in
distributed systems or cloud environments where physical and
administrative control is not guaranteed, a more robust system that
lacks this ``trusted arbiter'' requirement is desirable.

To overcome the need for a trusted enforcement mechanism in access
control systems, researchers have turned to cryptographically-based
access control systems. \cite{goyal2006}~and~\cite{bethencourt2007}
propose several cryptographically-based access control systems. These
systems are based on the concept of Attribute-Based Encryption
(ABE). ABE schemes allow a user to encrypt a document in a manner such
that the access control rules associated with the document are part of
the encryption process itself. Thus, in order to decrypt/access a
document, a user must satisfy one or more cryptographically guaranteed
access control attributes. \cite{goyal2006}~allows users to encrypt
documents that can only be decrypted by users possessing specific
attribute polices encoded in their keys.
\cite{bethencourt2007}~extends this concept to allow documents to be
encrypted with a full access control policy embedded in the encryption
itself, allowing only users who's private keys meet a generalized set
of requirements to access the documents. Both these systems allow the
construction of access control systems that do not require any trusted
arbiter to regulate access to objects. Instead, the access control
policy is enforced by the underlying cryptography itself.

Such concepts have not yet been widely deployed in day-to-day use,
possibly because of the complexity and computational overhead of
building and operating such systems. These systems also still push off
the generation, storage, and management of private keys to end users
and administrators, raising many of the same key-management challenges
discussed in Section~\ref{chap:challenges:solutions}.

\subsection{Homomorphic Encryption}

The rise of the cloud as the host to many modern data processing
systems has led to questions about the degree to which third party
cloud providers should be trusted with access to the data processed on
their infrastructure. Homomorphic encryption systems are designed to
help mitigate this trust. Such systems are designed to perform data
processing operations on encrypted data directly, avoiding the need to
decrypt it and expose the unencrypted data to a third party.

The previous ten years have heralded the arrival of numerous
partially-homomorphic encryption systems capable of performing certain
classes of data processing and manipulation operations directly on
encrypted data. System like CryptDB~\cite{popa2011} allow users to
search and query encrypted data directly, allowing the storage of such
databases on untrusted infrastructure. Other systems like
CS2~\cite{kamara2011} provided similar protections and capabilities in
more generic data storage contexts. Such systems are referred to as
partially homomorphic encryption schemes since they only allow a
subclass of all possible operations to be performed on encrypted
data. Such systems have been fielded and shown to be practical today.

Beyond partially homomorphic encryption schemes, fully homomorphic
encryption schemes allow for unrestricted classes of operations to be
performed on encrypted data~\cite{gentry2009}. A number of such
systems have been proposed -- some of which go so far as to propose
the possibility of encrypting entire computer programs that will be
executed by fully homomorphic processors~\cite{Breuer2013,
  Brenner2011}. Such system are certainly appealing, but thus far,
building one with low enough overhead to be practical for general
purpose usage has proven elusive.

While homomorphic solutions will likely prove to be part of the
solution to the problem of third party trust, they don't inherently
solve all trust-related problems. In the practical since, the
available homomorphic systems today are limited to only preforming
certain operations. Furthermore, while such systems allow the
processing of data atop untrusted infrastructure, they do not provide
a solution for use cases where you wish to share plain-text data with
other users or otherwise leverage an application that inherently
requires data to be decrypted. Homomorphic encryption systems also
risk leaking certain classes of data as a side effect of the
computation they perform. E.g. a homomorphic system may be able to
sort data without knowing its contents, but in doing so, it might
still learn something about the relative ordering of elements. As in
other cases, such systems also fail to provide a general solution for
the management of the associated cryptographic keys protecting such
operations.

\subsection{Secure Storage}

Beyond data processing, minimizing the degree of third party trust
required to securely store data atop cloud infrastructure is a
desirable goal. As such, a number of projects have undertaken efforts
aimed at achieving such protections. Secure data storage is one of the
most common class of security and privacy enhancing systems. And for
good reason -- the ability to securely store data is a critical
primitive for maintaining the security and privacy of computing
systems.

A number of traditional encrypted file systems exist with the design
goal of avoiding server-side trust~\cite{kher2005}. File systems like
CryptoCache~\cite{jensen2000}, RFS~\cite{dong2011}, and
Plutus~\cite{kallahalla2003} are all designed to allow users to store
files on servers without trusting the server itself. Other systems
like Keypad~\cite{geambasu2011} and CleanOS~\cite{tang2012} are aimed
at securing data atop user devices and protecting that data when a
device is lost or stolen. All of these systems employ various forms of
cryptography to obtain their goals.
 
Distributed storage systems like Depot~\cite{mahajan2011},
OceanStore~\cite{kubiatowicz2000}, and Tahoe~\cite{wilcox-o'hearn2008}
are all designed to minimize trust in the underlying storage
infrastructure. Such systems allow users to distribute and store files
across many third party nodes while also ensuring that the failure,
either accidental or intentional, of a subset of nodes does not result
in the loss, corruption, or exposure of user data. Depot focuses on
data integrity and availability in the presence of untrusted
nodes. Tahoe focus on data integrity and privacy in the presence of
untrusted nodes. OceanStore has elements of both as well as a focus on
large scale deployments that provide for properties like data locality
to enhance performance.

In many of these systems, however, key management primitives are still
ignored or pushed down to the user, leading to usability issues and
use-case mismatches. The SSaaS ideas proposed in this document could
be used to extend such designs to better account for the key
management challenges limiting the use of such systems today.

\section{Enhancing End User Security}
\label{chap:related:enduser}

The Edward Snowden leaks, as well as numerous highly publicized
privacy failures by companies ranging from Facebook to Target, have
fueled renewed calls for improved security and privacy enhancing
technologies aimed at end users. In response to these calls, a number
of organizations have begun offering new classes of security and
privacy enhancing tools. Many of these tools share the goals proposed
in this work -- namely, the creation of easy-to-use security and
privacy enhancing tools well adapted for modern user requirements.

\subsection{Communication Tools}

Many of the recent security and privacy enhancing tools have focused
on securing inter-user communication. Both the contents of and
meta-data associated with such communication has been the focus of
many of the recent NSA leaks~\cite{schneier-metadata}. The reaction to
such reports has spurred both the creation of new tools as well as the
expansion of the use of existing secure communion tools.

Email has long been the bedrock of digital communication mediums. The
traditional tools of for securing email, OpenPGP~\cite{openpgp} and
its various implementation (e.g. GnuPG~\cite{gnupg} and Symantec
PGP~\cite{pgp}), are not new. But they have seen renewed levels of
interest as effective mass-surveillance countermeasures. Unfortunately
these tools remain no more usable today then they were twenty years
ago, leading to the multitude of usability issues discussed in
Section~\ref{chap:background:usability}. In response, developer have
created tools like Mailpile~\cite{mailpile} and
Mailvelope~\cite{mailvelope} with the aim of making PGP-based email
security more friendly for the average user. Similarly, Google and
Yahoo are both engaging in efforts aimed and making PGP-like
encryption and authentication systems available to their webmail
users~\cite{google-endtoend, yahoo-endtoend}. Other systems such as
STEED aim to adapt traditional PGP principles to make the primitives
simpler for the average end user to deal
with~\cite{koch2011}. Finally, systems such as KeyBase~\cite{keybase}
aim to help users identify the public PGP keys of people with whom
they wish to communicate.\footnote{In some ways, KeyBase can be seen
  as the compliantly opposite of the SSaaS model discussed in this
  document. Whereas KeyBase aims to make it easy for users to discover
  and verify the public keys or other users, SSaaS aims to make it
  easy for users to protect their corresponding private keys.}

But even these recent pushes to re-skin PGP and make it more user
friendly can't overcome many of the fundamental issues with the
system~\cite{green-pgp}. Nor are they well suited for securing another
major form of modern communion: real-time chat. To overcome these
deficiencies, companies have released programs such as
TextSecure~\cite{openwhisper} and ChatSecure~\cite{chatsecure}. Both
systems rely on variants of the Off-the-Record (OTR) secure messaging
protocol to achieve encrypted, authenticated, and forward-secure
real-time communication between two or more parties~\cite{otr-v3,
  borisov2004, goldberg2009}. These systems aim to make secure
real-time communication as simple as possible, and have shown promise
in terms of popularity and usability. But these systems are not
necessarily direct replacements for PGP since they rely on the
real-time nature of chat communication to attain forward
secrecy. Similarly usable and forward-secure solutions for
non-real-time systems like email remain elusive.\footnote{Although,
  that has not stopped researcher from proposing some creative
  solutions to work around the forward secrecy problem, such as
  systems that force email to ``self-destruct'' after a certain
  age~\cite{geambasu2009}.}

While secure communication systems are seeing a lot of development and
showing promise, all of the solutions discussed above are
communication-specific solutions. As such, they are complimentary to
the work presented in this dissertation related to general methods for
reducing third party trust and managing secrets such as cryptographic
keys.

\subsection{Password Managers}

One of the most common classes of third party secret storage available
today is that of password storage by end user password managers. Due
to the well-documented failure of user-chosen
passwords~\cite{mazurek2013, goodin-bible, goodin-passwords,
  singer2013}, experts now recommend that users leverage password
management services. Such services rely on a single strong password to
protect a collection of unique and random passwords for each website
or service a user must access~\cite{schneier-passwords,
  krebs-passwords, brodkin-passman}.

Systems such as LastPass~\cite{lastpass} and
1Password~\cite{onepassword} provide users with a hosted platform on
which to manage and store their passwords. These systems utilize a
single third party for data storage in order to accommodate the
multi-device sync and multi-user sharing use cases presented in
Section~\ref{chap:challenges:usecases}. But in doing so, they require
the user to place a fair amount of trust in the third party
provider. Such systems do tend to perform client-side encryption and
generally are designed to avoid storing the user's master password in
a recoverable form, mitigating some of the potential for third party
abuse. But at the end of the day, such systems still require the user
to trust a single third party with their credentials, and by proxy,
access to their online services and accounts. Nor are such service
infallible. There have been several noted flaws and breaches of
password management systems over the previous few
years\footnote{Fortunately the deign of such systems has largely
  mitigated the worst damages from these
  breaches~\cite{ducklin2015}.}\cite{blackhat-lastpass,
  changedmy-lastpass, lastpass-blog-breach}.

Password management systems share some of the same goals proposed in
this document -- namely, the storage of end user secrets in a secure
and easy-to-use manner. Thus, password storage can be viewed as a
special case of the more generic Secret Storage as a Service (SSaaS)
model. This work presents more versatile and generalized solutions to
the secret storage problem. Such solution could be used to implement
the password manger concept while also decoupling such implementation
from needing to trust any single third party.

\subsection{Storage Tools}

Beyond secure communication and password storage lies the more general
problem of arbitrary secure data storage. As mentioned in
Section~\ref{chap:challenges:usecases}, most users expect such storage
to provide multi-user and multi-device support in addition to raw
secure storage, disqualifying many of the traditional encrypted file
system solutions from the running. Storage services like
Dropbox~\cite{dropbox} or Google Drive~\cite{google-drive} are
popular, but these solutions provide the associated third parties with
unfettered access to user data, making them unsuitable for users who
wish their data to remain private.

To overcome issues with systems like Dropbox while still affording
users access to modern file storage amenities, developers have creates
systems such as SpiderOak~\cite{spideroak}, Tresorit~\cite{tresorit},
and Wuala~\cite{wuala}. Such services aim to provide users with a
Dropbox-like alternative that does not require trusting the storage
provider. These services shift the encryption of data into the client,
helping to ensure the server only ever holds an encrypted copy of said
data. None the less, the confidentially and privacy claims made by
such services have been shown to be dubious~\cite{wilson2014}, largely
due to the fact the a user must still place a large degree of trust in
the backing third party in order to facilitate file sharing with other
users.

As in previous cases, however, these solutions are purpose built for
secure storage, and fail to provide a more general solution for
minimizing third party trust. Furthermore, as mentioned above, these
services still rely on a moderate degree of trust in a single third
party in order to facilitate sharing use cases. The work proposed in
this document could be leveraged to re-implement such systems while
further suppressing reliance on a single trusted third party.

\section{Key and Secret Management Systems}
\label{chap:related:keymgmt}

As mentioned in Section~\ref{chap:challenges:solutions}, one of the
main hindrances to bootstrapping secure systems is the lack of
versatile key-management systems capable of supporting modern use
cases while also minimizing third party trust. Such key management
systems are subsets of the larger secret storage problem discussed in
this proposal. This work, however, is not the first to recognize key
management as one of the critical components to building flexible and
secure privacy and security enhancing systems.

\subsection{Key Management in Storage Systems}

As discussed previously, many existing secure storage systems fail to
meet the needs of today's users due to the fact that their restrictive
and tightly coupled key management systems can't accommodate modern
use cases. A number of existing secure file systems have acknowledged
the issues created by tightly coupling key management with a specific
storage solution (or of ignoring key management all together and
forcing the user to deal with securing their keys themselves). These
systems represent a step toward more flexible key management, and by
proxy, more usable secure file storage.

Systems like Plutus~\cite{kallahalla2003} are designed to isolate key
management as a dedicated system component. While this is the first
step toward solving many of the key management problems, Plutus simply
shifts key management onto the client, providing some degree of
user-facing key management flexibility, but failing to provide a
standardized and general key management
system. SFS~\cite{mazieres1999} takes this separation a step further
by proposing various standardized mechanisms by which a user might
manage keys and an externally-facing interface for doing so. But even
SFS fails to define a general system for key management, instead
focusing on mechanisms that allow the user to select their own key
management system.

A system such as SFS, however, could be interfaced with a general
secret storage systems such as those proposed in this document. Such
applications of standardized secret storage systems are discussed
further in Chapter~\ref{chap:apps}.

\subsection{Key Escrow}

Key escrow systems represent on common form of key management system.
Key escrow systems aim to fit client-side key management into
hierarchical organizational frameworks where upper level members may
need access to subordinate's keys. Such systems have also been
proposed in regulatory environments where law enforcement or
regulatory personal must be able to access certain forms of otherwise
secure data. Escrow systems also prove beneficial in situations where
a key holder would like to hedge against the loss of their copy of the
key without having to fully trust any other single party with a backup
copy. Such systems often leverage secret-sharing schemes such as
Shamir~\cite{shamir1979} to accomplish their goals.

Early key escrow systems provided a trusted third party with a
secondary copy of a given data encryption key (and/or generated data
encryption keys in a manner such that two separate keys could be used
to decrypt any given piece of data)~\cite{denning1996}; e.g. the
infamous NSA-backed Clipper chip~\cite{whitehouse-clipper}. These
systems, however, fail the single-trusted-third party test, requiring
a high degree of user trust in one external entity. To overcome such
trust concentrations, researchers proposed distributed key escrow
systems that make key escrow requests transparent and leverages a
distributed consensuses to avoid any single rogue actor using such a
system to steal a key or gain unauthorized access to end user
data~\cite{blaze1996}.

While such distributed key management systems succeed in avoiding
trust in any single third party, they fail to provide for a more
general key management system that moves beyond traditional
escrow-based use cases (e.g. key sharing for multi-device sync or key
sharing for document collaboration). Furthermore, the existing
scholarship around key escrow systems fails to explore the benefits
and possibilities of the general purpose secret storage systems
described in this document. The secret storage systems described
herein could, however, be leveraged to fulfill many of the traditional
roles of a distributed key-escrow system.

\subsection{Automated and Cloud-based Secret Management}

The most closely related work to that proposed here is the various
automation and cloud-oriented key and secret management systems
introduced over the previous five years. These systems are a reaction
to many of the developer-facing use cases discussed in
Section~\ref{chap:challenges:usecases}. They aim to make key
management and secret storage primitives available to developers
leveraging automated configuration management systems and/or third
party cloud platforms.

Rackspace's CloudKeep~\cite{cloudkeep-presentation, cloudkeep} (now
OpenStack Barbican~\cite{openstack-barbican}) aims to create a
standardized key management system for use across multiple
applications, avoiding the need to re-implement such systems in each
application. Similar to the SSaaS system proposed in this document,
CloudKeep aims to ease developer burden while increasing the security
of end user applications by focusing security code in a centralized,
carefully curated system. CloudKeep/Barbican integrates with the
OpenStack~\cite{openstack} cloud infrastructure management platform to
provide secret storage services to OpenStack users. Such a system,
however, still requires the user to trust the OpenStack provider with
the storage of their secrets, and does not allow for the party
providing secret storage to be easily decoupled from the party
providing other cloud services.

Similarly, various IaaS platforms have begun offering cloud-based
Hardware Security Module (HSM) solutions. HSMs traditionally refer to
dedicated hardware co-processors capable of storing user cryptographic
keys in a manner that prevents them from ever being extracted. HSMs
are then tasked with performing all necessary cryptographic operations
requiring such keys on the user's behalf. Standardized interfaces such
as PKCS11~\cite{pcks11-standard} exist to allow programs to
communicate and utilize such hardware. The benefits of such chips lie
in their ability to securely store cryptographic keys and to perform
cryptographic operations without exposing such keys to shared computer
memory space (and potentially leaking them in the process, as occurred
in Heartbleed~\cite{heartbleed}). Amazon has recently begun offering a
cloud-backed HSM service that aims to make the benefits of dedicated
HSM hardware available in the virtualized world of the
cloud~\cite{amazon-hsm}. Such systems, while helpful for developers
who wish to build HSM-dependent systems that operate on local or cloud
hardware, still require placing a high degree to trust in a single
third party to faithfully operate virtual HSMs. The SSaaS work
discussed in this proposal could potentially be used to build similar
``soft'' HSM systems without requiring single third party
trust~\cite{lorier-pkcs11}.

Beyond infrastructure-tied services, a number of open source secret
storage systems have been developed by industry, including HashiCorp's
Vault~\cite{vault}, Lyft's Confidant~\cite{confidant}, and Square's
Keywhiz~\cite{keywhiz}. Similar commercial systems also exist with the
aim of providing developers with turnkey key-management
systems~\cite{gazzang, porticor, rosen2012}. The development of and
demand for such systems demonstrate the clear challenges industry
faces managing secrets. Such systems, however, still generally require
placing trust in a single third party, or in the case of self-hosted
solutions, a single server.\footnote{Although this party may now
  differ from the party providing the underlying cloud compute or data
  storage servers.} They also lack the cross-platform standardization
necessary to incentive some of market-driven forms of trust-preserving
behavior discussed in Chapters~\ref{chap:trust} and~\ref{chap:ssaas}.

%%  LocalWords:  CryptoCache OceanStore SFS Plutus Custos SSaaS TPP
%%  LocalWords:  Rackspace's CloudKeep Custos's CryptDB CleanOS NSA
%%  LocalWords:  Snowden Mailpile TextSecure ChatSecure LastPass IaaS
%%  LocalWords:  OnePassword Diffie SpiderOak Tresorit Wuala Shamir
%%  LocalWords:  HSMs PKCS Heartbleed Snowden's OpenStack Barbican
%%  LocalWords:  Mailvelope HashiCorp's Lyft's Keywhiz OTR KeyBase
%%  LocalWords:  keybase
