% 
% PhD Dissertation Thesis
%
% 2016
%
% Andy Sayler
% www.andysayler.com


\documentclass[twoside,openright,defaultstyle,11pt]{thesis}
%\documentclass[defaultstyle,11pt]{thesis}

% System Packages
\usepackage{array}
\usepackage{epsfig}
\usepackage[dvipsnames]{xcolor}
\usepackage{hyperref}
\usepackage{url}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tabu}
\usepackage{tikz-qtree}
\usepackage{fancyvrb}
\usepackage{listings}
\usepackage{emptypage}

% Package Options
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}
\captionsetup{justification=centering}
\lstset{
  language={},
  basicstyle=\footnotesize,
  numbers=left,
  numberstyle=\tiny,
  stepnumber=1,
  numbersep=5pt,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=4,
  captionpos=b,
  breaklines=true,
  breakatwhitespace=false,
  frame=single,
  frameround=tttt
}
\lstdefinelanguage{JavaScript}{
  keywords={break, case, catch, continue, debugger, default, delete,
    do, else, finally, for, function, if, in, instanceof, new, return,
    switch, this, throw, try, typeof, var, void, while, with, true,
    false, null},
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]',
  morestring=[b]'',
  sensitive=true
}

% Macros
\input macros.tex

% Other Options
\clubpenalty = 10000
\widowpenalty = 10000

%%%%%%%%%%%%   All the preamble material:   %%%%%%%%%%%%

\title{Securing Secrets and Managing Trust in Modern Computing Applications}

\author{Andy}{Sayler}

\otherdegrees{B.S.E.E., Tufts University, 2011 \\
              M.S.C.S., University of Colorado, 2013}

\degree{Doctor of Philosophy}   %  #1 {long descr.}
	   {Dissertation}           %  #2 {short descr.}

\dept{Department of}		    %  #1 {designation}
     {Computer Science}         %  #2 {name}

\advisor{Prof.}                 %  #1 {title}
        {Dirk Grunwald}         %  #2 {name}

\reader{Prof. Eric Keller}      %  2nd person to sign thesis
\readerThree{Prof. John Black}  %  3rd person to sign thesis
\readerFour{Prof. Sangtae Ha}   %  4th person to sign thesis
\readerFive{Prof. Blake Reid}   %  5th person to sign thesis

\abstract{

  \OnePageChapter

  The amount of digital data generated and stored by users increases
  every day. In order to protect this data, modern computing systems
  employ numerous cryptographic and access control solutions. Almost
  all of such solutions, however, require the keeping of certain
  secrets as the basis of their security models. How best to securely
  store and control access to these secrets is a significant
  challenge: such secrets must be stored in a manner that protects
  them from a variety of potentially malicious actors while still
  enabling the kinds of functionality users expect.

  This dissertation discusses a system for isolating secrets from the
  applications that rely on them and storing these secrets via a
  standardized, service-oriented secret storage system.  This ``Secret
  Storage as a Service'' (SSaaS) model allows users to reduce the
  trust they must place in any single actor while still providing
  mechanisms to support a range of cloud-based, multi-user, and
  multi-device use cases.

  This dissertation contains the following contributions: an overview
  of the secret-storage problem and how it relates to the security and
  privacy of modern computing systems and users, a framework for
  evaluating the degree by which one must trust various actors across
  a range of popular use cases and the mechanisms by which this trust
  can be violated, a description of the SSaaS model and how it helps
  avoid such trust and security failures, a discussion of how the
  SSaaS approach can integrate with and improve the security of a
  range of applications, an overview of Custos -- a first-generation
  SSaaS prototype, an overview of Tutamen -- a next-generation SSaaS
  prototypes, and an exploration of the legal and policy implications
  of the SSaaS ecosystem.

}

\dedication[Dedication] { Dedicated to Edward Snowden, Chelsea
  Manning, Bill Binney, Thomas Drake, Perry Fellwock, Russ Tice, Mark
  Klein, Thomas Tamm, and all those who have risked and who will risk
  their lives and livelihoods to expose the privacy and security
  abuses of governments around the world. }

\acknowledgements{
  \OnePageChapter

  This document could not have been completed without the assistance
  of a variety of individuals. First, thanks to all of my committee
  members for their oversight, suggestions, and time. Dirk Grunwald,
  my advisor, provided numerous suggestions and coauthored the Custos
  and Tutamen papers that predated this document. Eric Keller provided
  suggestions and assistance on a variety of topics including the
  original Custos implementation that was undertaken as project for
  his course. Blake Reid provided the impetus for me to get more
  involved in the policy side of technology and was kind enough to
  have me as a member of the Colorado Technology Law and Policy Clinic
  (TLPC) during the Spring of 2015. John Black and Sangtae Ha have
  both provided suggestions and answers a variety of technical
  questions.

  Beyond those that served on my committee, many other individuals
  also provided technical support and insights. Matt Monaco provided
  the implementation of the Tutamen \texttt{dm-crypt} FDE client and
  is a coauthor on the Tutamen paper. Taylor Andrews provided the
  Tutamen-backed encrypted Dropbox client implementation and is also a
  coauthor on the Tutamen paper. Joseph Lorenzo Hall and Erik Stallman
  both provided mentorship and guidance on a variety of security and
  policy topics during my time in Washington, DC at the Center for
  Democracy and Technology.

  Finally, a big thanks to my partner Denali Hussin for her continual
  support throughout my time as a PhD student, as well as her editing
  support on a range of (overly) verbose technical documents. Thanks
  also to my parents Mike and Lori for their $27+$ years of support
  and assistance. And thanks to all those who assisted and supported
  my efforts, but whom I have unintentionally failed to mention here.

}

% \ToCisShort % use this only for 1-page Table of Contents

\LoTisShort   % use this only for 1-page Table of Tables
% \emptyLoT   % use this if there is no List of Tables

% \LoFisShort % use this only for 1-page Table of Figures
% \emptyLoF	  % use this if there is no List of Figures

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%       BEGIN DOCUMENT...         %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\input chap_intro.tex
\input chap_background.tex
\input chap_challenges.tex
\input chap_related.tex
\input chap_trust.tex
\input chap_ssaas.tex
\input chap_apps.tex
\input chap_custos.tex
\input chap_tutamen.tex
\input chap_policy.tex
\input chap_conclusion.tex

%%%%%%%%%   then the Bibliography, if any   %%%%%%%%%
\bibliographystyle{plain} % or "siam", or "alpha", etc.
\nocite{*}                % list all refs in database, cited or not
\bibliography{refs}       % Bib database in "refs.bib"

%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
%\appendix
%\singlespacing

\end{document}

%%  LocalWords:  SSaaS tex Custos ssaas custos tutamen Sangtae FDE
%%  LocalWords:  dmcrypt Hussin TLPC Stallman
