% 
% PhD Dissertation Thesis
%
% 2016
%
% Andy Sayler
% www.andysayler.com


\documentclass[defaultstyle,11pt]{thesis}

% System Packages
\usepackage{array}
\usepackage{epsfig}
\usepackage{color}
\usepackage{hyperref}
\usepackage{url}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tabu}
\usepackage{tikz-qtree}
\usepackage{fancyvrb}
\usepackage{listings}

% Package Options
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}
\lstset{
  language={},
  basicstyle=\footnotesize,
  numbers=left,
  numberstyle=\tiny,
  stepnumber=1,
  numbersep=5pt,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=4,
  captionpos=b,
  breaklines=true,
  breakatwhitespace=false,
  frame=single,
  frameround=tttt
}
\lstdefinelanguage{JavaScript}{
  keywords={break, case, catch, continue, debugger, default, delete,
    do, else, finally, for, function, if, in, instanceof, new, return,
    switch, this, throw, try, typeof, var, void, while, with, true,
    false, null},
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]',
  morestring=[b]'',
  sensitive=true
}

% Macros
\input macros.tex

% Other Options
\clubpenalty = 10000
\widowpenalty = 10000

%%%%%%%%%%%%   All the preamble material:   %%%%%%%%%%%%

\title{Securing Secrets and Managing Trust in Modern Computing Applications}

\author{Andy}{Sayler}

\otherdegrees{B.S.E.E., Tufts University, 2011 \\
              M.S.C.S., University of Colorado, 2013}

\degree{Doctor of Philosophy}   %  #1 {long descr.}
	   {Dissertation}           %  #2 {short descr.}

\dept{Department of}		    %  #1 {designation}
     {Computer Science}         %  #2 {name}

\advisor{Prof.}                 %  #1 {title}
        {Dirk Grunwald}         %  #2 {name}

\reader{Prof. John Black}       %  2nd person to sign thesis
\readerThree{Prof. Sangtae Ha}  %  3rd person to sign thesis
\readerFour{Prof. Eric Keller}  %  4th person to sign thesis
\readerFive{Prof. Blake Reid}   %  5th person to sign thesis

\abstract{

  \OnePageChapter

  The amount of digital data generated and stored by users increases
  every day. In order to protect this data, modern computing systems
  employ numerous cryptographic and access control solutions. Almost
  all of such solution, however, require the keeping of certain
  secrets as the basis of their security models. How best to store,
  secure, and control access to such secrets is a significant
  challenge. Such secrets must be stored in a manner that secures them
  from a variety of potentially malicious devices and actors while
  also enabling the kinds of functionality users expect today.

  In order to facilitate such flexible but secure storage of secrets,
  I propose isolating secrets from the applications that rely on them
  and storing these secrets via a standardized, service-oriented
  secret storage system.  This ``Secret Storage as a Service'' (SSaaS)
  model allows users to reduce the trust they must place in any single
  device or actor while still providing mechanisms to support the
  range of cloud-based, multi-user, multi-device use-cases popular
  today.

  In this document, I begin by presenting an overview of the
  secret-storage problem and how it relates to the security and
  privacy of user data today. I continue by introducing a framework
  for evaluating the degree to which we must trust various systems and
  parties across a range of popular usage patterns, as well as the
  mechanisms by which this trust can be violated. I then discuss ways
  in which the SSaaS model helps avoid such trust and security
  failures. I go on to discuss the qualities a desirable SSaaS
  ecosystem would have as well as some of the previously proposed
  secret storage solutions. I then present both first-generation and
  next-generation SSaaS prototypes and discuss the improvements they
  make over previous secret storage systems. Next, I discuss how an
  SSaaS approach would integrate with a range of applications,
  including the presentation of several prototype SSaaS-backed
  applications. Finally I discusses the legal and policy implications
  of such a system.

}

\dedication[Dedication]{
  Todo
}

\acknowledgements{
  \OnePageChapter
  Todo
}

%\ToCisShort % use this only for 1-page Table of Contents

\LoFisShort	% use this only for 1-page Table of Figures
% \emptyLoF	% use this if there is no List of Figures

\LoTisShort	% use this only for 1-page Table of Tables
% \emptyLoT	% use this if there is no List of Tables

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%       BEGIN DOCUMENT...         %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\input chap_intro.tex
\input chap_background.tex
\input chap_challenges.tex
\input chap_related.tex
\input chap_trust.tex
\input chap_ssaas.tex
\input chap_custos.tex
\input chap_tutamen.tex
\input chap_apps.tex
%\input chap_planned.tex

%%%%%%%%%   then the Bibliography, if any   %%%%%%%%%
\bibliographystyle{plain} % or "siam", or "alpha", etc.
\nocite{*}                % list all refs in database, cited or not
\bibliography{refs}       % Bib database in "refs.bib"

%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
%\appendix
%\singlespacing

\end{document}

%%  LocalWords:  SSaaS tex Custos ssaas custos
